Monitoring Time Series With Missing Values:\\a Deep Probabilistic Approach

### 2022-02-17 paper.tex
first submittable version

>>>>

We evaluate the proposed architecture on a data set of
monitoring a cluster of 100 computing nodes in the cloud. For
each node, the incoming and the outgoing network traffic (in
bytes) and the CPU usage (relative) are logged with 1 minute
resolution. 240 hours were logged, resulting in 12000 120-minute
3-dimensional samples.  We split the dataset into the training,
validation, and test as 80\%, 10\%, and 10\% correspondingly.
Since the original data set does not have any missing data
points, we emulated data sets with missing data by randomly
removing 5\%, 10\%, 20\%, and 50\% of the data.

We used a 3-layer GRU-based recurrent neural network with hidden
size 64 and 20\% dropout between layers. We trained the network
with lookahead depths (number of steps to forecast in the
future) 2, 4, 8, and 16 using the Adam optimizer with learning
rate 0.001, training for 20 epochs (sufficient for convergence).
We performed the training on a cloud computing node with 1
NVIDIA T4 GPU, 4 Intel Xeon Platinum CPUs, and 64 Gb memory. The
training of a single model took 20 minutes.

We compared our approach with conventional imputation methods
`replace by the mean' and `replace by a random sample'.
In the `replace by mean' method, a missing value is replaced 
by the mean of the forecast. In the `replace by a random
sample' method, a missing value is replaced by a random sample
drawn from the forecast. As a performance metrics, we
used per-point negative log-likelihood loss on the test set. 
Tables~\ref{tab:replace-by-mean} and~\ref{tab:replace-by-sample}
show the difference in loss between uncertainty propagation and
`replace by the mean' and `replace by a random sample',
correspondingly. The greater is the number, the worse is the
forecasting by each of the methods compared to uncertainty
propagation.
\begin{table}[t]
\caption{Uncertainty propagation vs. `replace by the mean`.}
\label{tab:replace-by-mean}
	\setlength\tabcolsep{18pt}
	\centering
	\begin{tabular}{r|c c c c}
	 \textbf{missing} & \textbf{2} & \textbf{4} & \textbf{8} & \textbf{16} \\ \hline
		  \textbf{5\%} & 0.001 & 0.001 &  0.06 & 0.10  \\
		 \textbf{10\%} & 0.001 & 0.002 &  0.08 & 0.11  \\
		 \textbf{20\%} & 0.003 & 0.003 &  0.11 & 0.13  \\
		 \textbf{50\%} & 0.004 & 0.006 &  0.12 & 0.16  
	\end{tabular}
\end{table}
\begin{table}[t]
\caption{Uncertainty propagation vs. `replace by a random sample`.}
\vspace{1em}
\label{tab:replace-by-sample}
	\setlength\tabcolsep{18pt}
	\centering
	\begin{tabular}{r|c c c c}
	 \textbf{missing} & \textbf{2} & \textbf{4} & \textbf{8} & \textbf{16} \\ \hline
		  \textbf{5\%} & 0.04 & 0.04 & 0.13 & 0.14 \\
		 \textbf{10\%} & 0.06 & 0.06 & 0.16 & 0.24 \\
		 \textbf{20\%} & 0.11 & 0.12 & 0.20 & 0.27 \\
		 \textbf{50\%} & 0.18 & 0.19 & 0.28 & 0.30   
	\end{tabular}
\end{table}
One can see that in all cases uncertainty propagation provides
better forecasts than either of the conventional methods.

The code and data for the case studies are available at
\url{https://bitbucket.org/dtolpin/dbts-studies/}.
